Predictive Modeling of I/O Characteristics \\ in High Performance Computing Systems

### 2018-03-08 Paper_[2017-12-15]_HPC/paper.tex
Changed some folder names, made final modifications to papers for submission.

%     Methodology     
%=====================
\section{Methodology}
\label{sec:methodology}
\subsection{Data}
\vspace{-10pt}
In order to evaluate the viability of multivariate models for
predicting system performance, this paper presents a case study of a
four-dimensional dataset produced by executing the IOzone benchmark
from \shortciteN{iozone} on a homogeneous cluster of computers. The
system performance data was collected by executing IOzone 40 times for
each of a select set of system configurations. A single IOzone
execution reports the max I/O throughput seen for the selected
test. The 40 executions for each system configuration are converted
into the mean and variance, both values in $\mathbb{R}$ capable of
being modeled individually by the multivariate approximation
techniques presented in Section \ref{sec:multivariate}. The summary of
data used in the experiments for this paper can be seen in Table
\ref{tab:data_type}.  Distributions of raw throughput values being
modeled can be seen in Figure \ref{fig:raw_throughput}.

@@

A na\"{\i}ve multivariate prediction technique such as nearest
neighbor could experience relative errors in the range $\displaystyle
[0, \big(\max_x f(x) - \min_x f(x)\big) / \min_x f(x) ]$ when modeling
a nonnegative function $f(x)$ from data. The IOzone mean data response
values span three orders of magnitude (as can be seen in Table
\ref{tab:data_type}) while variance data response values span six
orders of magnitude. It is expected therefore, that all studied
multivariate models perform better than a na\"{\i}ve approach,
achieving relative errors strictly less than $10^3$ for throughput
mean and $10^6$ for throughput variance. Ideally, models will
yield relative errors significantly smaller than 1.

@@

The present results demonstrate that a straightforward application of
multivariate modeling techniques can be used to effectively predict
HPC system performance. Some modeling effort on the part of a systems
engineer combined with a significant amount of experimentation (days
of CPU time for the IOzone data used here) can yield a model capable
of accurately tuning an HPC system to the desired performance
specification, although qualitatively correct predictions can be
achieved with mush less (10\%, say) effort.

@@

>>>>

%     Methodology     
%=====================
\section{Methodology}
\label{sec:methodology}
\subsection{Data}
\vspace{-10pt}

In order to evaluate the viability of multivariate models for
predicting system performance, this paper presents a case study of a
four-dimensional dataset produced by executing the IOzone benchmark
from \shortciteN{iozone} on a homogeneous cluster of computers. All
experiments were performed on parallel shared-memory nodes common to
HPC systems. Each system had a lone guest Linux operating system
(Ubuntu 14.04 LTS//XEN 4.0) on a dedicated 2TB HDD on a 2 socket, 4
core (2 hyperthreads per core) Intel Xeon E5-2623 v3 (Haswell)
platform with 32 GB DDR4. The system performance data was collected by
executing IOzone 40 times for each of a select set of system
configurations. A single IOzone execution reports the max I/O
throughput seen for the selected test in kilobytes per second. The 40
executions for each system configuration are converted into the mean
and variance, both values in $\mathbb{R}$ capable of being modeled
individually by the multivariate approximation techniques presented in
Section \ref{sec:multivariate}. The summary of data used in the
experiments for this paper can be seen in Table \ref{tab:data_type}.
Distributions of raw throughput values being modeled can be seen in
Figure \ref{fig:raw_throughput}.

@@

A na\"{\i}ve multivariate prediction technique such as nearest
neighbor could experience relative errors in the range $\displaystyle
[0, \big(\max_x f(x) - \min_x f(x)\big) / \min_x f(x) ]$ when modeling
a nonnegative function $f(x)$ from data. The IOzone mean data response
values span three orders of magnitude (as can be seen in Table
\ref{tab:data_type}) while variance data response values span six
orders of magnitude. It is expected therefore, that all studied
multivariate models perform better than a na\"{\i}ve approach,
achieving relative errors strictly less than $10^3$ for throughput
mean and $10^6$ for throughput variance. Ideally, models will yield
relative errors significantly smaller than 1. The time required to
compute thousands of models involved in processing the IOzone data
through MDA was approximately five hours on a CentOS workstation with
an Intel i7-3770 CPU at 3.40GHz. In four dimensions for example, each
of the models could be constructed and evaluated over hundreds of
points in less than a few seconds.

@@

The present results demonstrate that a straightforward application of
multivariate modeling techniques can be used to effectively predict
HPC system performance. Some modeling effort on the part of a systems
engineer combined with a significant amount of experimentation (days
of CPU time for the IOzone data used here) can yield a model capable
of accurately tuning an HPC system to the desired performance
specification, although qualitatively correct predictions can be
achieved with much less (10\%, say) effort.

@@

%     Author Biography Block     
%================================
\section*{Author Biographies}

\textbf{\uppercase{Thomas C. H. Lux}} is a Ph.D. student at Virginia Tech
studying computer science under Dr. Layne Watson.

\textbf{\uppercase{Layne T. Watson}} (Ph.D., Michigan, 1974) has interests
in numerical analysis, mathematical programming, bioinformatics, and data
science.
He has been involved with the organization of HPCS since 2000.

\textbf{\uppercase{Tyler H. Chang}} is a Ph.D. student at Virginia Tech
studying computer science under Dr. Layne Watson.

\textbf{\uppercase{Jon Bernard}} is a Ph.D. student at Virginia Tech
studying computer science under Dr. Kirk Cameron.

\textbf{\uppercase{Bo Li}} is a senior Ph.D. student at Virginia Tech
studying computer science under Dr. Kirk Cameron.

\textbf{\uppercase{Li Xu}} is a Ph.D. student at Virginia Tech studying
statistics under Dr. Yili Hong.

\textbf{\uppercase{Godmar Back}} (Ph.D., University of Utah, 2002) has
broad interests in computer systems, with a focus on performance and
reliability aspects of operating systems and virtual machines.

\textbf{\uppercase{Ali R. Butt}} (Ph.D., Purdue, 2006) has interests in cloud computing, distributed computing, and operating system induced variability.

\textbf{\uppercase{Kirk W. Cameron}} (Ph.D., Louisiana State, 2000) has
interests in computer systems design, performance analysis, and power and
energy efficiency.

\textbf{\uppercase{Yili Hong}} (Ph.D., Iowa State, 2009) has interests in
engineering statistics, statistical modeling, and data analysis.
